TIM: Tree Imaging Machine for low-cost digital images of tree cores and cookies

### 2025-03-27 docs/writing/paper.tex
Addressing Lizzies Mar 9 edits.

% !TeX root = paper.tex
\documentclass[a4paper,12pt]{article}

% Packages
\usepackage[utf8]{inputenc}     
\usepackage{amsmath, amssymb} 
\usepackage{graphicx}          
\usepackage{geometry}           
\usepackage[numbers]{natbib}            
\usepackage{hyperref}          
\usepackage{subcaption}     
% Page layout
\geometry{margin=1in}

@@

\begin{abstract}  %emw9Mar: I believe dendro is tree ring DATING, so I adjusted below to be more inclusive. 
Studies of tree rings from climatology to ecology have benefitted recently from new methods to image wood rapidly and efficiently, helping to scale up the number and quality of image samples. Current open-source approaches are, however, limited in dimensionality and can be cost-prohibitive. The Tree Imaging Machine (TIM) is a do-it-yourself scanning tool made to digitize tree cookies and cores that address these limitations. TIM takes partially overlapping microscopic images of samples and stitches the individual images together to form 
a mosaic, which can be zoomed in to visualize features on the scale of 0.01 mm. With scans of up to 21 140 DPI, TIM produces one of the highest resolution images among similar tools for the competitive price of less than \$3 000 USD.
We designed TIM to have a large working plane to allow digitizing cores over 50 cm and complete cookies. Operators can prepare sample batches of multiple cores and cookies, letting the machine work until the queue is finished.
All of the 3D printed parts are available for download on the [NIH 3D printing repository](https://doi.org/10.60705/3DPX/21561.3) and the software is 
open source, with instructions to recreate this tool available on [GitHub](https://github.com/temporalecologylab/TreeRings). TIM provides an important advance in scanning resolution, tree cookie sampling, maximum core sample size, batch digitization, and cost in a single package. This promotes the digitization of larger samples and can collect data at finer resolutions in an open-source design which can be built on and improved as new technologies become available. % Add to the discussion something about next steps for TIM more clearly? 
\end{abstract}

@@

% I would start a new paragraph here that opens with something like: Since XX was introduced in XX, image capture and stitching of tree rings has increased. Today CaptuRING offers a  more modern do-it-yourself alternative to ATRICS that is (open source? Has resolutions up to XX? Used XX type of camera? Give a little more here). Another approach, Gigapixel, moves the camera relative to the sample, allowing for multiple samples to be recorded in sequence \citep{griffin_gigapixel_2021}. 
% Then I would end on the cookies issue and also anything else to mention, like cost? You should also clearly mention when/how stitching happens in these approaches.
Since ATRICS was introduced in 2007, image capture and stitching of tree rings have become more common. Today CaptuRING offers a more modern do-it-yourself alternative to ATRICS, creating scans up to 5 339 dpi using a DSLR camera and motorized stage \citep{garcia-hidalgo_capturing_2022}. 
Another alternative, Gigapixel, achieves an impressive 19 812 dpi with the same DSLR image capture and composite stitch method \citep{griffin_gigapixel_2021}. 
While CaptuRING and Gigapixel both use DSLR cameras and motorized movement, they differ on two fronts. 
First, Gigapixel's camera is traversed across a stationary sample while CaptuRING moves a core sample underneath a stationary camera.
With a mobile camera, multiple samples can be recorded sequentially without needing to remove a sample from a stage upon image completion---significantly reducing the active setup time for an operator. 
Second, Gigapixel automatically stitches individual images together once imaging is complete. This removes the need to transfer images to a more powerful computer for stitching as is necessary with CaptuRING.
Additionally, both of these systems are designed to capture images along a singular axis as the entire width of a core fits within a single image's field of view.
This assumption fundamentally restricts the digitization to be limited only to cores, as cookies have a two dimensional scanning surface. %emw9Mar: later 2/3 of paragraph very nice.

% I will edit this a little next time, but for now fix up em-dashes, and simplify with less PV as possible. 
We made TIM to combine the open-source and open-hardware philosophies of CaptuRING with the functionality of Gigapixel.
We added support to digitize both cookies and cores, increased the maximum sample length, perform image stitching directly without the need for manual file transfer, and capture batches of samples sequentially---all while minimizing cost.
Our design requires the use of 3D printed components and common hand tools such as allen keys and wrenches. No specialized equipment or power tools are required beyond this.
Excluding 3D printed parts, the total cost of the machine is approximately \$2 200 USD \citep[compared to the \$70 000 USD of the Gigapixel,][]{griffin_gigapixel_2021}.
We chose to use a Raspberry Pi HQ camera with a microscope lens to save on purchasing a professional camera with a macro lens. Professional camera setups are comparable in price to the entire cost of TIM.
Additional savings can also be had when factoring in the cost of a professional stitching software license such as PTGui, the suggested software to use with CaptuRING.

@@

% Good topic sentence! Can you walk reader more carefully through each item? You may need to re-order them .. But basically, pick the most useful order and then write: For the camera .... For the gantry machine ... You might also use this space to say, in contrast to previous approaches that have used XX giant camera, we used ... and otherwise point out the novelty when it comes up. 
TIM can be thought of as a combination of multiple subsystems: a gantry machine, camera, and computer (See Figure \ref{fig:tim_assembled}). 
For the gantry machine, cartesian movement in the $X$, $Y$ and $Z$ directions is a result of two machine kits and a motor controller from OpenBuilds---the ACRO 1010, the NEMA 17 lead screw linear actuator, and the X32 Motor Controller running GRBL firmware. 
The machine kit includes extruded aluminum rails, carriages to slide along the rails, stepper motors to control the $X$ and $Y$ movement, and all the hardware to assemble the machine. 
To fit on the build plates of the ACRO system, we made an adapter to connect the linear actuator to the X and Y axis, thus creating motion in the Z axis.
The size of the work plane is ultimately up to the size of the ACRO kit that is purchased---allowing for more flexibility.
Building on top of kits allowed for quick assembly with sound instructions and saved development time. 
For the camera, we chose a 12MP Raspberry Pi HQ Camera equipped with a SEEED studio microscope lens and connected this to the gantry machine using custom adapters.
For the computer, we implemented all of the software control, image stitching, and graphical user interface (GUI) to run on an NVIDIA Jetson Orin Nano (Jetson) which also connects to the camera via a MIPI cable.
Controlling the machine is done by launching a Python program and interacting with mouse button clicks on the GUI.

%emw9Mar: check my edits are accurate, I think having a reason beyond getting a nicer computer would be good and is true to our decision making. 
By choosing this combination, we were able to reduce the weight and cost of the camera significantly, which reduced design challenges in stabilizing movement of a heavier DSLR and allowed us to invest more in a powerful computer that can handle intensive image processing. 
The Jetson is an edge computer which drives a monitor for a GUI, sends commands to the motor controller to move the machine, runs image processing calculations for control, and performs calculations to stitch individual images into one mosaic. 
Despite the weight reductions, torsion on the gantry arm still resulted in a non-zero torsional deflection. A torsion correcting adapter was designed to counteract this rotation and level the lens.

\subsection{Preparing Cookies and Cores}
% Tried to reduce PV below. 
TIM can scan both tree cookies and tree cores. Sample preparation for both are equally important but require different approaches.
The general guidelines used for preparing cookies and cores applies---requiring samples to be sanded with incrementally increasing sand paper grit to produce an even surface for scanning \citep{speer_fundamentals_2010}. We sanded cookies using an orbital sander with mesh sandpaper from 60 to 800 grit, then used a microscope to visually inspect the quality of the sanding. We considered the sample to be sufficiently prepared when vessels were easily identifiable.

%emw9Mar: You have to number figures in the order they are introduced, but I think you could introduce Fig 2 below under 'nearly parallel' and then you don't need to re-order. 
TIM requires additional sanding considerations for each sample type. 
For cookies, it is important to have the top and bottom surfaces nearly parallel. Small differences in plane angle can be corrected using the 3D printed leveling table we designed (See Figure \ref{fig:ideal_levelling}). 
For cores, care should be taken to maximize the scanning surface. This means removing material to be coincident to the center of the cross section of the core. 
Similarly to cookies, the scanning surface should be as close to parallel to the XY-plane as possible, but generally does not need to be leveled using the leveling tables.
Cores also need to be aligned to be parallel to the $Y$ axis to be digitized, using squaring tool or similar to help with alignment. Any warping in core mounts needs to be counteracted with the use of trigger clamps.

@@

In the GUI the operator indicates the height and width of the field of view in an individual image from the camera. 
This varies on the focal length of the lens, but we have kept the field of view to be 3 mm x 5 mm. 
Once prompted the system begins to exhaustively traverse the surface area of the sample, using the sample height and width entered by the operator.  
The goal of this traversal is to obtain in-focus images that have a region of overlap with all its neighbors---the basis of image stitching (See Figure \ref{fig:cookie_traverse}). 
Digitizing cores can be done without needing to traverse in both the X and Y axis. 
The optimal scanning path for a core is to align the length of cores to the Y axis and only take one column of images.
Spanning the surface area of a cookie requires movement in both the X and Y axis.

@@

\begin{figure}
  \centering
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[height=0.5\linewidth]{../diagrams/sample_setup_ideal.png}
    \caption{Ideal sample leveling}
    \label{fig:ideal_levelling}
  \end{subfigure}%
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[height=0.5\linewidth]{../diagrams/sample_setup_realistic.png}
    \caption{Realistic sample leveling}
    \label{fig:realistic_levelling}
  \end{subfigure}
  \caption{Side view of the camera and sample on top of a leveling table. The ideal sample leveling shows a uniform distance d at all $(X,Y)$ coordinates on the sample. This is impossible to achieve in reality, the true sample leveling has a non uniform distance at unique $(X_i, Y_i)$.}
  \label{fig:sample_levelling}
\end{figure}

@@

\[
  i_{\max} = \arg\max_{i} NV(image(\boldsymbol{{Z_\text{focus}})})
  \]

The second stage of our time-efficient focusing procedure is using a feedback loop to update the initial $Z_0$ at each image location. 
Without a feedback loop, the sample alignment must have a difference in height no greater than 0.5 mm, $d_1 - d_2 < 0.5 mm $ (See Figure \ref{fig:realistic_levelling}). 
Rather than adjusting this range, a greater alignment error can be managed by controlling the center of the range, $Z_{0,k}$, for $(X_k,Y_k)$. 
The likelihood of an adjacent %emw9Mar:  we need a noun here instead of just $(X_{k+1}, Y_{k+1})$ ... maybe we need a term for this?
$(X_{k+1}, Y_{k+1})$ containing an in-focus image is highest when the current $i_{{\max,k}}$ is at the middle index of $\boldsymbol{Z_{samples}}$. 
A PID control algorithm with a process variable of $i_{max}$ and control variable of $Z_{0,k}$ allows a negative feedback loop to improve focusing across the entire sample \citep{odwyer_summary_2000}. 
Instead of forcing a remarkably level sample alignment of $(d_1 - d_2) < 0.5mm$ (See Figure \ref{fig:realistic_levelling}), the system can handle $(Z_{k} - Z_{k+1} < 0.5mm)$ which is readily achievable. %After the capturing overlapping images across the entire surface area, the images are ready to be stitched.

@@

\section{Results} %emw9Mar: I wondered about this being Results & Discussion ... would MEE be okay with that? It would work better for the paper I think. What do the guidelines say on formatting for articles?

@@

\begin{figure}
    \centering
    \includegraphics[height=0.5\linewidth]{../../code/plots/time_and_area.png}
    \caption{Time to digitize a sample is dependent on the sample's surface area and the desired final resolution.
    Cores benefit from linear surface area to sample length. The range of sample included are from a 3mm x 220mm core to a 75mm x 56mm cookie. True sampling times are drastically influenced by configurable machine parameters.} % Take the title on top off this image as we don't have titles on images in sci papers. 
    \label{fig:digitization_time}
\end{figure}

\section{Discussion} % I think we could remind the reader here of what gaps TIM addresses, but I can help do that next time.  %emw9Mar: This section could move down and become conclusions if we merge the Results & Discussion. 
By taking advantage of a common three degree of freedom cartesian machine design, powerful edge computing, and a microscope camera we were able to design a cost effective and high resolution digitization tool for wood samples. 
We significantly increased the maximum sample length and maximum resolution compared to common alternatives in the field. 
In addition, the machine design was intended to be readily replicated by labs with minimal engineering experience and equipment.

\subsection{Strengths and Opportunities} %emw9Mar: And this section could become part of the Results & Discussion.... 
% add better opportunities
TIM's design minimizes the barriers to build an affordable, efficient and high-resolution image capture system in smaller labs without compromising scan quality, and while allowing scanning of cookies and cores. We have measured tree ring widths using the digitized samples from TIM using CooRecorder as well as ImageJ.
Extensions to TIM could allow automatic uploads of images into cloud data storage system with automatic uploads, though this would likely require significant effort for it to work across diverse cloud platforms. 
% removing the need to manually transfer completed scans with physical drives. %emw9Mar: But we just do this because of our wifi issues ... others could manually upload them to the cloud, right? So I suggest we remove this. 
%emw9Mar: Any citations for developing repos to share tree ring images? Or just more on building better open shared data for tree rings? We should also cite some of the new vessel and ring ID papers in CITES2
One opportunity for this development could come as more universal and large-scale open-data storage of tree ring images come online CITES. Such open repositories would also increase the value of developing new frameworks count vessel or identify rings, which are already starting and have the potential to greatly decrease the time invested in manually identifying tree anatomy CITES2. With these more fully developed TIM could be extended to be run such frameworks while scanning, possibly taking additional scans only when needed to reach sufficient algorithm accuracy. %emw9Mar -- check my riffing here.

Our implementation of focusing images could be improved with the aid of additional sensors. Much of the imaging time of samples is the result of taking a vertical stack of images to obtain one in focus image. 
In advanced microscopy applications, it is common to see supplemental sensors used to measure the distance between the subject and lens.
Adding sensor support and a control loop to focus images could significantly reduce sampling time on the scale of an order of magnitude. 
Using rotary encoders on each of the motors controlling the X and Y movement would be a welcomed addition---this would greatly increase the 
accuracy of the the location of the camera on the work plane.  
Finally, an exciting frontier for machines such as TIM is integrating additional sensors. While much of the current imaging time of samples comes from taking a vertical stack of images to obtain one in-focus image, advanced microscopy applications have overcome similar challenges through supplemental sensors that measure the distance between the subject and lens.
Adding such sensor support and a control loop to focus images could significantly reduce sampling time on the scale of an order of magnitude.

>>>>

% !TeX root = ./paper.tex
\documentclass[a4paper,12pt]{article}

% Packages
\usepackage[utf8]{inputenc}     
\usepackage{amsmath, amssymb} 
\usepackage{graphicx}          
\usepackage{geometry}           
\usepackage[numbers]{natbib}            
\usepackage{hyperref}          
\usepackage{subcaption}     
\usepackage{multirow}

% Page layout
\geometry{margin=1in}

@@

\begin{abstract}  
Studies of tree rings from climatology to ecology have benefitted recently from new methods to image wood rapidly and efficiently, helping to scale up the number and quality of image samples. Current approaches are, however, limited in scanning dimensionality and can be cost-prohibitive. % Getting rid of open-source as CaptuRING is cheap but GigaPixel is expensive and not open source
The Tree Imaging Machine (TIM) is a do-it-yourself scanning tool made to digitize tree cookies and cores that address these limitations. TIM takes partially overlapping microscopic images of samples and stitches the individual images together to form 
a mosaic, which can be zoomed in to visualize features on the scale of 0.01 mm. With scans of up to 21 140 DPI, TIM produces one of the highest resolution images among similar tools for the competitive price of less than \$3 000 USD.
We designed TIM to have a large working plane to allow digitizing cores over 50 cm and complete cookies. Operators can prepare sample batches of multiple cores and cookies, letting the machine work until the queue is finished.
All of the 3D printed parts are available for download on the [NIH 3D printing repository](https://doi.org/10.60705/3DPX/21561.3) and the software is  % is this best practice for including links? 
open source, with instructions to recreate this tool available on [GitHub](https://github.com/temporalecologylab/TreeRings). TIM provides an important advance in scanning resolution, tree cookie sampling, maximum core sample size, batch digitization, and cost in a single package. This promotes the digitization of larger samples and can collect data at finer resolutions in an open-source design which can be built on and improved as new technologies become available. % Archive GitHub with Zenodo? I'm nervous as the code base is updated relatively frequently... will this be a problem? Or do I only ever make one 'release' https://docs.github.com/en/repositories/archiving-a-github-repository/referencing-and-citing-content 
\end{abstract}

@@

Since ATRICS was introduced in 2007, image capture and stitching of tree rings have become more common. Today CaptuRING offers a more modern do-it-yourself alternative to ATRICS, creating scans up to 5 339 dpi using a DSLR camera and motorized stage \citep{garcia-hidalgo_capturing_2022}. 
Another alternative, Gigapixel, achieves an impressive 19 812 dpi with the same DSLR image capture and composite stitch method \citep{griffin_gigapixel_2021}. 
While CaptuRING and Gigapixel both use DSLR cameras and motorized movement, they differ on two fronts. 
First, Gigapixel's camera is traversed across a stationary sample while CaptuRING moves a core sample underneath a stationary camera.
With a mobile camera, multiple samples can be recorded sequentially without needing to remove a sample from a stage upon image completion---significantly reducing the active setup time for an operator. 
Second, Gigapixel automatically stitches individual images together once imaging is complete. This removes the need to transfer images to a more powerful computer for stitching as is necessary with CaptuRING.
Additionally, both of these systems are designed to capture images along a singular axis as the entire width of a core fits within a single image's field of view.
This assumption fundamentally restricts the digitization to be limited only to cores, as cookies have a two dimensional scanning surface. %emw9Mar: later 2/3 of paragraph very nice.

We made TIM to combine the open-source and open-hardware philosophies of CaptuRING with the functionality of Gigapixel.
We added support to digitize both cookies and cores, increased the maximum sample length, perform image stitching directly without the need for manual file transfer, and capture batches of samples sequentially---all while minimizing cost.
Our design requires the use of 3D printed components and common hand tools such as allen keys and wrenches. No specialized equipment or power tools are required beyond this.
Excluding 3D printed parts, the total cost of the machine is approximately \$2 200 USD \citep[compared to the \$70 000 USD of the Gigapixel,][]{griffin_gigapixel_2021}.
We chose to use a Raspberry Pi HQ camera with a microscope lens to save on purchasing a professional camera with a macro lens. Professional camera setups are comparable in price to the entire cost of TIM.
Additional savings can also be had when factoring in the cost of a professional stitching software license such as PTGui, the suggested software to use with CaptuRING.

@@

% Good topic sentence! Can you walk reader more carefully through each item? You may need to re-order them .. But basically, pick the most useful order and then write: For the camera .... For the gantry machine ... You might also use this space to say, in contrast to previous approaches that have used XX giant camera, we used ... and otherwise point out the novelty when it comes up. 
TIM can be thought of as a combination of multiple subsystems: a gantry machine, camera, and computer (Figure \ref{fig:tim_assembled}). 
For the gantry machine, cartesian movement in the $X$, $Y$ and $Z$ directions is a result of two machine kits and a motor controller from OpenBuilds---the ACRO 1010, the NEMA 17 lead screw linear actuator, and the X32 Motor Controller running GRBL firmware. 
The machine kit includes extruded aluminum rails, carriages to slide along the rails, stepper motors to control the $X$ and $Y$ movement, and all the hardware to assemble the machine. 
To fit on the build plates of the ACRO system, we made an adapter to connect the linear actuator to the X and Y axis, thus creating motion in the Z axis.
The size of the work plane is ultimately up to the size of the ACRO kit that is purchased---allowing for more flexibility.
Building on top of kits allowed for quick assembly with sound instructions and saved development time. 
For the camera, we chose a 12MP Raspberry Pi HQ Camera equipped with a SEEED studio microscope lens and connected this to the gantry machine using custom adapters.
For the computer, we implemented all of the software control, image stitching, and graphical user interface (GUI) to run on an NVIDIA Jetson Orin Nano (Jetson) which also connects to the camera via a MIPI cable.
Controlling the machine is done by launching a Python program and interacting with mouse button clicks on the GUI.

By choosing this combination, we were able to reduce the weight and cost of the camera significantly, which reduced design challenges in stabilizing movement of a heavier DSLR and allowed us to invest more in a powerful computer that can handle intensive image processing. 
The Jetson is an edge computer which drives a monitor for a GUI, sends commands to the motor controller to move the machine, runs image processing calculations for control, and performs calculations to stitch individual images into one mosaic. 
Despite the weight reductions, torsion on the gantry arm still resulted in a non-zero torsional deflection. A torsion correcting adapter was designed to counteract this rotation and level the lens.

\subsection{Preparing Cookies and Cores}
TIM can scan both tree cookies and tree cores. Sample preparation for both are equally important but require different approaches.
The general guidelines used for preparing cookies and cores applies---requiring samples to be sanded with incrementally increasing sand paper grit to produce an even surface for scanning \citep{speer_fundamentals_2010}. We sanded cookies using an orbital sander with mesh sandpaper from 60 to 800 grit, then used a microscope to visually inspect the quality of the sanding. We considered the sample to be sufficiently prepared when vessels were easily identifiable.

TIM requires additional sanding considerations for each sample type. 
For cookies, it is important to have the top and bottom surfaces nearly parallel. 
\begin{figure}
  \centering
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[height=0.5\linewidth]{../diagrams/sample_setup_ideal.png}
    \caption{Ideal sample leveling}
    \label{fig:ideal_levelling}
  \end{subfigure}%
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[height=0.5\linewidth]{../diagrams/sample_setup_realistic.png}
    \caption{Realistic sample leveling}
    \label{fig:realistic_levelling}
  \end{subfigure}
  \caption{Side view of the camera and sample on top of a leveling table. The ideal sample leveling shows a uniform distance d at all $(X,Y)$ coordinates on the sample. This is impossible to achieve in reality, the true sample leveling has a non uniform distance at unique $(X_i, Y_i)$.}
  \label{fig:sample_levelling}
\end{figure}
Small differences in plane angle can be corrected using the 3D printed leveling table we designed (Figure \ref{fig:ideal_levelling}). 
For cores, care should be taken to maximize the scanning surface. This means removing material to be coincident to the center of the cross section of the core. 
Similarly to cookies, the scanning surface should be as close to parallel to the XY-plane as possible, but generally does not need to be leveled using the leveling tables.
Cores also need to be aligned to be parallel to the $Y$ axis to be digitized, using squaring tool or similar to help with alignment. Any warping in core mounts needs to be counteracted with the use of trigger clamps.

@@

In the GUI, the operator indicates the height and width of the field of view in an individual image from the camera. 
This varies on the focal length of the lens, but we have kept the field of view to be 3 mm x 5 mm. 
Once prompted the system begins to exhaustively traverse the surface area of the sample, using the sample height and width entered by the operator.  
The goal of this traversal is to obtain in-focus images that have a region of overlap with all its neighbors---the basis of image stitching (Figure \ref{FIGURE LABEL}). 
Digitizing cores can be done without needing to traverse in both the X and Y axis. 
The optimal scanning path for a core is to align the length of cores to the Y axis and only take one column of images.
Spanning the surface area of a cookie requires movement in both the X and Y axis.

@@

@@

\[
  i_{\max} = \arg\max_{i} NV(image(\boldsymbol{{Z_\text{samples}})})
  \]

The second stage of our time-efficient focusing procedure is using a feedback loop to update the initial $Z_0$ at each image location. 
Without a feedback loop, the sample alignment must have a difference in height no greater than 0.5 mm, $d_1 - d_2 < 0.5 mm $ (Figure \ref{fig:realistic_levelling}). 
Rather than adjusting this range, a greater alignment error can be managed by controlling the center of the range, $Z_{0,k}$, for $(X_k,Y_k)$. 
The likelihood of an adjacent %emw9Mar:  we need a noun here instead of just $(X_{k+1}, Y_{k+1})$ ... maybe we need a term for this?
$(X_{k+1}, Y_{k+1})$ containing an in-focus image is highest when the current $i_{{\max,k}}$ is at the middle index of $\boldsymbol{Z_{samples}}$. 
A PID control algorithm with a process variable of $i_{max}$ and control variable of $Z_{0,k}$ allows a negative feedback loop to improve focusing across the entire sample \citep{odwyer_summary_2000}. 
Instead of forcing a remarkably level sample alignment of $(d_1 - d_2) < 0.5mm$ (Figure \ref{fig:realistic_levelling}), the system can handle $(Z_{k} - Z_{k+1} < 0.5mm)$ which is readily achievable. %After the capturing overlapping images across the entire surface area, the images are ready to be stitched.

@@

\section{Results \& Discussion} 
%emw9Mar: I wondered about this being Results & Discussion ... would MEE be okay with that? It would work better for the paper I think. What do the guidelines say on formatting for articles? 
% MEE's guidelines have a distinct Results section from Discussion although maybe they're flexible on this? Conclusion is optional for MEE

@@

\begin{figure}
    \centering
    \includegraphics[height=0.5\linewidth]{../../code/plots/time_and_area.png}
    \caption{Time to digitize a sample is dependent on the sample's surface area and the desired final resolution.
    Cores benefit from linear surface area to sample length. The range of sample included are from a 3mm x 220mm core to a 75mm x 56mm cookie. True sampling times are drastically influenced by configurable machine parameters.} 
    \label{fig:digitization_time}
\end{figure}

\subsection{Strengths and Opportunities} %emw9Mar: And this section could become part of the Results & Discussion.... 
% add better opportunities
TIM's design minimizes the barriers to build an affordable, efficient and high-resolution image capture system in smaller labs without compromising scan quality, and while allowing scanning of cookies and cores. We have measured tree ring widths using the digitized samples from TIM using CooRecorder as well as ImageJ.
%emw9Mar: Any citations for developing repos to share tree ring images? Or just more on building better open shared data for tree rings? 
% It appears databases store data after images are processed and measured, potentially too expensive to store image datasets at high resolution?
Publicly available tree-ring datasets can be found on multiple online repositories such as TreeSource's Tree-Rings Database for Canadian sites or the International Tree-Ring Data Bank (ITRDB) for worldwide data \citep{girardin_national_2021} \citep{grissino-mayer_international_1997}.
Both of these databases offer metadata rich data on measured tree-rings but do not store the source images of samples.
One opportunity for this development could come as more universal and large-scale open-data storage of tree ring images come online. 
Such open repositories would also increase the value of developing new frameworks count vessel or identify rings, which are already starting and have the potential to greatly decrease the time invested in manually identifying tree anatomy \citep{resente_mask_2021} \citep{polacek_automation_2023}. 
With these more fully developed TIM could be extended to be run such frameworks while scanning, possibly taking additional scans only when needed to reach sufficient algorithm accuracy. %emw9Mar -- check my riffing here.

% Sandy got back to me! :)
We tested a scanned cookie from TIM with the R-CNN presented by \citeauthor{polacek_automation_2023} on two servers. The first being the high-performance UBC ARC Sockeye which completed the task in 15 minutes. As many do not have access to such powerful computing, we also tested
the same image on an RTX A6000 GPU server---taking 3.5 hours to complete the task for approximately \$3.50 USD. 
While the results of automated tree rings are promising, further testing is necessary to test the feasibility of running this model on an edge computer or developing integrations for TIM with a remote server.

Finally, an exciting frontier for machines such as TIM is integrating additional sensors. While much of the current imaging time of samples comes from taking a vertical stack of images to obtain one in-focus image, advanced microscopy applications have overcome similar challenges through supplemental sensors that measure the distance between the subject and lens.
Adding such sensor support and a control loop to focus images could significantly reduce sampling time on the scale of an order of magnitude. 
Using rotary encoders on each of the motors controlling the X and Y movement would be a welcomed addition---this would greatly increase the 
accuracy of the the location of the camera on the work plane.

\section{Conclusion} % I think we could remind the reader here of what gaps TIM addresses, but I can help do that next time.  %emw9Mar: This section could move down and become conclusions if we merge the Results & Discussion. 
By taking advantage of a common three degree of freedom cartesian machine design, powerful edge computing, and a microscope camera we were able to design a cost effective and high resolution digitization tool for wood samples. 
We significantly increased the maximum sample length and maximum resolution compared to common alternatives in the field. 
In addition, the machine design was intended to be readily replicated by labs with minimal engineering experience and equipment.

\subsection{Supplemental Information}

\begin{tabular}{ |p{3cm}||p{12cm}|  }
  \hline
  \multicolumn{2}{|c|}{Math Notation} \\
  \hline
  Terms & Definition\\
  \hline
  $X_0, Y_0, Z_0$   &The initial cartesian coordinates of the camera when the sample is saved by the user. \\
  $(X_k, Y_k, Z_k)$ & A unique set of $X$, $Y$, and $Z$ coordinates for the center point of an individual image $k$. \\
  $(X_{k+1}, Y_{k+1})$ & The $X$ and $Y$ coordinates of an image that neighbors image $k$ at $(X_k, Y_k)$. \\
  $Z_{0,i}$ & The starting $Z$ value to move to when moving to the adjacent image. This value is also the center value of $\boldsymbol{Z_{\text{samples}}}$ \\
  $NV()$ & Normalized variance 'function.' \\
  $Z_{focused}$ & The $Z$ value of the most in focus image in $\boldsymbol{Z_{\text{samples}}}$\\
  $\boldsymbol{Z_{\text{samples}}}$ & A vector containing the $Z$ coordinate for multiple images at the same $X_k, Y_k$. \\
  $i_{\max,k}$ & The index of the image with the largest normalized variance at the same $(X_k, Y_k)$. \\
  $w$ & The width of the window to traverse when trying to center a core. \\
  $image()$ & The image at a corresponding $Z$ value from $\boldsymbol{Z_{\text{samples}}}$ \\
  $d_1$ & The relative distance from the camera to one edge of a levelled sample. \\
  $d_2$ & The relative distance from the camera to the opposite edge of a levelled sample. \\

  \hline
 \end{tabular}
